Sample Length at Which Block Bootstrapping is Effective for Estimation of Parameter of Time Series

### 2023-02-10 Manuscript/manuscript.tex
opening paragraph

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}

@@

\begin{abstract}
Block bootstrapping is a method that is useful for estimating a parameter of a time
series. Theoretically, the method will work perfectly given an infinitely large sample of the time series.
It is necessary to know how large of a finite sample is required to for block bootstrap estimation to work. Before this study, there have been no investigations into this problem in the literature. The aim of this study is to answer this question by simulating replications of block bootstrap interval estimations of a parameter of a time series, and recording the rate at which the parameter is recovered by each replicate interval. The sample length required for acceptable performance is unsurprisingly found to be dependent on the type of interval construction and the series' level of temporal dependence.

@@

@@

Block bootstrapping is a widely used method in Statistics. The concept for block 
bootstrapping was developed independently by \citet{hall1985resampling}, \citet{carlstein1986use}, and 
\citet{kunsch1989jackknife}. \citet{radovanov2014comparison} It has been applied to a variety of 
different domains such 
as econometrics \citep{mackinnon2006bootstrap} and meteorology \citep{varga2017generalised}. It can be used for 
situations in which there is temporal dependence within a population, and the goal is to estimate or test a hypothesis about a parameter. Assuming that the sample is infinitely 
large, the method should work perfectly. However, for a finite sample size, the method may 
not work as well as expected. In the context of planning an applied 
statistical procedure, for which a large sample size is not always practical, it may be helpful to know how large the sample must be for the method to perform at an acceptable level.

\citet{hesterberg2015teachers} notes that while percentile-based confidence intervals from basic nonparametric bootstrapping (simple resampling with no blocks involved) are more accurate than t-intervals for larger sample sizes, they are less accurate for smaller sample sizes. \citet{nevitt2001performance} find that a sample size of 200-1000 is usually sufficient for interval estimation using basic nonparametric bootstrapping.
\citet{goncalves2005bootstrap} found that standard error estimates from block bootstrapping small 
samples may be significantly more accurate than inference from closed-form asymptotic estimates. That study focused on the estimation of a coefficient parameter within the context of linear regression. Still, the block bootstrap percentile confidence intervals under-covered the parameter even for n = 1024.

@@

%At the base level, we are block-bootstrapping an auto-regressive process with true mean 0, and some AR(1) coefficient.

%Bootstrapping is the procedure of creating new pseudo-samples of the same size by resampling observations with replacement from the original sample. Basic bootstrapping works well for samples
% that are do not have a dependence factor. However, for a time series such as an auto-regressive or moving-average process, a different procedure is required to account for the time dependence. In such a situation, instead of resampling observations, blocks of the original sample can be resampled (these blocks must be of a certain length by which the sample length is divisible) to create a pseudo-sample. Point estimates of a parameter calculated from many block bootstrapped pseudo-samples are then used to create a distribution to create an interval estimation of the parameter.

@@

%The various ways which these intervals can be constructed are described in \citet{diciccio1996bootstrap} and \citet{rice2006mathematical}. They will be discussed later in a review of the block bootstrap method. For the purposes of this study (in which the true parameters are known), in order to establish whether or not the method is effective, the block bootstrapping of a simulated sample can be replicated many times, and the proportion of confidence intervals that recover the true known parameter (coverage rate) is recorded. The goal of this is simply to observe how successful the block bootstrapping procedure is at estimating a parameter. The key variable being observed was, the sample length (n). It is known that as n increases, block bootstrapping should become a more accurate and precise method for estimation(the coverage rates will approach the theoretical confidence level and the interval widths will decrease). The question is what is the minimum sample length necessary to reach a coverage rate of 1 - $\alpha$/2 \%.

@@

%In order to solve the question of what sample length is necessary for effective block bootstrapping while still accounting for all these factors, the simulations were repeated for different variations of the MA coefficient, target parameter, and non-moving vs moving method. For each of these variations, results for block bootstrap estimation of mean, MA coefficient, and standard deviation were recorded for a wide range of sample length.

@@

Block bootstrapping is a method that can be used to estimate a parameter of serially dependent data. Suppose that one wants to estimate a parameter or parameters $\theta$ of a population. The first step is to generate a sample of length $n$ from the population. An estimate of the parameter $\hat{\theta}_{n}$ can be computed from this sample. A pseudo-estimate $\hat\theta_n^*$ can be computed from a bootstrapped pseudo-sample. The mean $\bar\theta_n^*$ of many replicate $\hat\theta_n^*$ can also be computed. Lastly, for each $\hat\theta_n^*$, $\delta^* = \hat\theta_n^* - \bar\theta_n^*$ can be computed.

>>>>

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref, url}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}

@@

\begin{abstract}
Block bootstrap is a method that is useful for estimating a parameter of a time
series. Theoretically, the method will work perfectly given an infinitely large sample of the time series.
It is necessary to know how large of a finite sample is required to for block bootstrap estimation to work. Before this study, there have been no investigations into this problem in the literature. The aim of this study is to answer this question by simulating replications of block bootstrap interval estimations of a parameter of a time series, and recording the rate at which the parameter is recovered by each replicate interval. The sample length required for acceptable performance is unsurprisingly found to be dependent on the type of interval construction and the series' level of temporal dependence.

@@

\jy{Practice my writing tips on latex (e.g., keep line width under 80; use
  double blank lines to separate paragraphs; etc.):
\url{https://statds.github.io/stat-writing/using-the-right-tools-for-writing.html}}

@@

Block bootstrap is an important tool for constructing confidence intervals when
making inferences about dependent data. Early ideas were developed independently
by \citet{hall1985resampling}, \citet{carlstein1986use}, and 
\citet{kunsch1989jackknife}. % \citet{radovanov2014comparison}
It has been applied to a variety of different domains such 
as econometrics \citep{mackinnon2006bootstrap} and meteorology
\citep{varga2017generalised}. It is particularly useful for serially dependent
data when the serial dependence is not specified or of not primary interest.
As the sample size goes to infinity, block bootstrap confidence intervals are
expected to have coverage rates matching their nominal levels. For finite sample
sizes, however, a natural question is, how large a sample size is large enough
for block bootstrap confidence intervals to have the desired coverage rates?

\citet{hesterberg2015teachers} notes that while percentile-based confidence intervals from basic nonparametric bootstrap (simple resampling with no blocks involved) are more accurate than t-intervals for larger sample sizes, they are less accurate for smaller sample sizes. \citet{nevitt2001performance} find that a sample size of 200-1000 is usually sufficient for interval estimation using basic nonparametric bootstrap.
\citet{goncalves2005bootstrap} found that standard error estimates from block bootstrap small 
samples may be significantly more accurate than inference from closed-form asymptotic estimates. That study focused on the estimation of a coefficient parameter within the context of linear regression. Still, the block bootstrap percentile confidence intervals under-covered the parameter even for n = 1024.

@@

%At the base level, we are block-bootstrap an auto-regressive process with true mean 0, and some AR(1) coefficient.

%Bootstrap is the procedure of creating new pseudo-samples of the same size by resampling observations with replacement from the original sample. Basic bootstrap works well for samples
% that are do not have a dependence factor. However, for a time series such as an auto-regressive or moving-average process, a different procedure is required to account for the time dependence. In such a situation, instead of resampling observations, blocks of the original sample can be resampled (these blocks must be of a certain length by which the sample length is divisible) to create a pseudo-sample. Point estimates of a parameter calculated from many block bootstrapped pseudo-samples are then used to create a distribution to create an interval estimation of the parameter.

@@

%The various ways which these intervals can be constructed are described in \citet{diciccio1996bootstrap} and \citet{rice2006mathematical}. They will be discussed later in a review of the block bootstrap method. For the purposes of this study (in which the true parameters are known), in order to establish whether or not the method is effective, the block bootstrap of a simulated sample can be replicated many times, and the proportion of confidence intervals that recover the true known parameter (coverage rate) is recorded. The goal of this is simply to observe how successful the block bootstrap procedure is at estimating a parameter. The key variable being observed was, the sample length (n). It is known that as n increases, block bootstrap should become a more accurate and precise method for estimation(the coverage rates will approach the theoretical confidence level and the interval widths will decrease). The question is what is the minimum sample length necessary to reach a coverage rate of 1 - $\alpha$/2 \%.

@@

%In order to solve the question of what sample length is necessary for effective block bootstrap while still accounting for all these factors, the simulations were repeated for different variations of the MA coefficient, target parameter, and non-moving vs moving method. For each of these variations, results for block bootstrap estimation of mean, MA coefficient, and standard deviation were recorded for a wide range of sample length.

@@

Block bootstrap is a method that can be used to estimate a parameter of serially dependent data. Suppose that one wants to estimate a parameter or parameters $\theta$ of a population. The first step is to generate a sample of length $n$ from the population. An estimate of the parameter $\hat{\theta}_{n}$ can be computed from this sample. A pseudo-estimate $\hat\theta_n^*$ can be computed from a bootstrapped pseudo-sample. The mean $\bar\theta_n^*$ of many replicate $\hat\theta_n^*$ can also be computed. Lastly, for each $\hat\theta_n^*$, $\delta^* = \hat\theta_n^* - \bar\theta_n^*$ can be computed.
